>

### 2024-12-04 docs/abstract/_abstract-full.qmd
Incorporate suggestion from Spyros

**Background:**
Networks have been used in different scientific fields to model complex systems. In medicine and epidemiology, drug prescription networks (DPN) have been proposed as a modeling tool for studying the temporal dynamics of medication distribution and consumption within a population.

**Objectives:**  
Evaluate the structural properties of temporal DPNs that are constructed from the observed daily co-prescriptions between treatments for mental health problems and other drug therapies in a pharmacy dispensing database.

>>>>

**Background:**
Networks have been used in different scientific fields to model complex systems. In medicine and epidemiology, drug prescription networks (DPN) have been proposed as a modeling tool for studying the temporal dynamics of medication co-prescription within a population.

**Objectives:**  
This study aims to assess the structural characteristics of temporal DPNs, derived from daily co-prescriptions between treatments for mental health conditions and other therapeutic drug classes, as recorded in pharmacy dispensing database.

### 2024-12-04 docs/article.qmd
Incorporate suggestion from Spyros

The relational characteristics in health research variables have fueled the growing implementation of network analysis, making it an indispensable research framework in medicine and epidemiology [@Luke2007]. Past implementations of network analysis in public health mostly focused on transmission network, social network, and organizational network. Only recently, @Cavallo2012 demonstrated a novel framework to describe medication uses on a population level, i.e. drug prescription network (DPN), a term which later is coined by @Bazzoni2015. Network analysis and graph theory offer an approach to model the relationship among building blocks of a complex system. From a complex system standpoint, polypharmacy is a prescription of multiple drug combinations which can be modelled as a network [@Miglio2021]. DPN captures a complex and dynamic system that may vary overtime, which is highly relevant to drug prescribed for mental disorders.

@@

Monitoring polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Shrivastava2013], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy monitoring system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. To adequately address the complexity, network analysis is a viable framework for analyzing polypharmacy. Transforming prescription registry into a DPN has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

@@

@@

@@

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] is equal to one, a network with similarly connected nodes will have an eigenvector centrality $c_i$ = $\frac{1}{n}$ for each node, where $n$ is the number of nodes. The value of $c_i = \frac{1}{n}$ is the expected eigenvector centrality in a densely connected network where connection $\mathcal{V}_{i, j}$ from node $\mathcal{V}_i$ to node $\mathcal{V}_j$ has the same probability for every node. Per the central limit theorem, the value $\frac{1}{n}$ can be parameterized as a mean of a normal distribution, where all instances of eigenvector centrality will follow the distribution, $c_i \sim \mathcal{N}(\mu, \sigma)$. A one-sample Student's T-test was performed to assess how much each eigenvector centrality differs from the expected values. As such, eigenvector centrality greater than $\frac{1}{n}$ can be considered high, and the value below it can be considered low. The p-values was corrected for a type-I error using Bonferroni method.

>>>>

The inherently complex nature of health-related phenomena have fueled the growing implementation of network analysis, making it an indispensable research framework in medicine and epidemiology [@Luke2007]. Past implementations of network analysis in public health mostly focused on transmission network, social network, and organizational network. Only recently, @Cavallo2012 demonstrated a novel framework to describe medication uses on a population level, i.e. drug prescription network (DPN), a term which later is coined by @Bazzoni2015. Network analysis and graph theory offer an approach to model the relationship among building blocks of a complex system. From a complex system standpoint, polypharmacy is a prescription of multiple drug combinations which can be modelled as a network [@Miglio2021]. DPN can capture the complex relationships of a system that may vary overtime, which is highly relevant to drug prescribed for mental disorders.

@@

Monitoring polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Shrivastava2013], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy monitoring system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. Network analysis is a viable framework for analyzing polypharmacy. Constructing DPNs from drug prescription data has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

@@

## Graph theory

- What characterizes a graph?
- When a graph is directed/undirected?
- Intuition behind an adjacency matrix

@@

## Centrality measures in graph theory

- Discuss about centrality measures in social network analysis
- What can centrality measures offer in analyzing polypharmacy networks
- Check the structure of complex network: theory and application by Estrada

@@

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] is equal to one, a network with similarly connected nodes will have an eigenvector centrality $c_i$ = $\frac{1}{n}$ for each node, where $n$ is the number of nodes. The value of $c_i = \frac{1}{n}$ is the expected eigenvector centrality in a densely connected network where connection $(i, j) \in \mathcal{E}$ from node $\mathcal{V}_i$ to node $\mathcal{V}_j$ has the same probability for every node. Per the central limit theorem, the value $\frac{1}{n}$ can be parameterized as a mean of a normal distribution, where all instances of eigenvector centrality will follow the distribution, $c_i \sim \mathcal{N}(\mu, \sigma)$. A one-sample Student's T-test was performed to assess how much each eigenvector centrality differs from the expected values. As such, eigenvector centrality greater than $\frac{1}{n}$ can be considered high, and the value below it can be considered low. The p-values was corrected for a type-I error using Bonferroni method.

### 2024-12-04 docs/exec-summary/_spectral.qmd
Draft an initial report using Elsevier template
Incorporate suggestion from Spyros

### Decomposition with singular spectrum analysis

Classical and STL-based decomposition technique may not sufficiently capture complex periodic patterns in a time-series. Singular spectrum analysis (SSA) is a powerful non-parametric technique by leveraging Hankel matrix as a higher-dimensional embedding of the time-series. The resulting Hankel matrix $X$ has the size of $L \times K$, where $L$ represents the lag term, i.e. a window length to perform embedding For a weekly data with a hypothesized yearly seasonality, the $L$ is set as 52. In theory, the number of $L$ should be between the range of $2 \leq L \leq \frac{N}{2}$, where $N$ is the total length of the time-series.

$$
X =
\stackrel{\textrm{Time-series embedding}}{
  \begin{bmatrix}
  x_1 & x_2 & \dots & x_{N - L + 1} \\
  x_2 & x_3 & \dots & x_{N - L + 2} \\
  \vdots & \vdots & \ddots & \vdots \\
  x_L & x_{L + 1} & \dots & x_{N} \\
  \end{bmatrix}
}
$$ {#eq-hankel-mtx}

The embedded time-series $X$ is then decomposed using a singular value decomposition (SVD). With a matrix $S = X \cdot X^T$, we can extract an eigenvalue $\lambda \ni \{\lambda_1 \geq \dots \geq \lambda_L\}$, where each $\lambda_i$ is a non-negative integer. Similarly, we can define $U$ as the eigenvectors of matrix $S$ corresponding to eigenvalues $\lambda$. Afterwards, we can extract factor vectors $V$ where the corresponding $V_i = X^T \frac{U_i}{\sqrt{\lambda_i}}$. Finally, the eigentriple of $\{\sqrt{\lambda_i}, U_i, V_i^T\}$ is formulated as a row-wise decomposition of matrix $X$ [@Golyandina2020]. The resulting eigentriple is then used to reconstruct the time-series by eigentriple grouping and diagonal averaging.

>>>>

::: {.content-visible unless-meta="in.article"}
### Decomposition with singular spectrum analysis
:::

Classical and STL-based decomposition technique may not sufficiently capture complex periodic patterns in a time-series. Singular spectrum analysis (SSA) is a powerful non-parametric technique by leveraging Hankel matrix as a higher-dimensional embedding of the time-series. The resulting Hankel matrix $X$ has the size of $L \times K$, where $L$ represents the lag term. The lag term is the length of data point in a time-series being taken to construct the higher-dimensional embedding. For a weekly data with a hypothesized yearly seasonality, the $L$ is set as 52, representing the number of week in a year. In theory, the number of $L$ should be between the range of $2 \leq L \leq \frac{N}{2}$, where $N$ is the total length of the time-series [@Golyandina2020].

$$
\tau =
\stackrel{\textrm{Time-series data}}{
  \begin{bmatrix}
  x_1 & x_2 & \cdots & x_{N-1} & x_N \\
  \end{bmatrix}
}
\quad \to \quad
X =
\stackrel{\textrm{Time-series embedding}}{
  \begin{bmatrix}
  x_1 & x_2 & \dots & x_{N - L + 1} \\
  x_2 & x_3 & \dots & x_{N - L + 2} \\
  \vdots & \vdots & \ddots & \vdots \\
  x_L & x_{L + 1} & \dots & x_{N} \\
  \end{bmatrix}
}
$$ {#eq-hankel-mtx}

The embedded time-series $X$ is then decomposed using a singular value decomposition (SVD). With a matrix $S = X \cdot X^T$, we can extract an eigenvalue $\lambda \ni \{\lambda_1 \geq \dots \geq \lambda_L\}$, where each $\lambda_i$ is a non-negative integer. Similarly, we can define $U \ni \{U_1, \dots, U_L\}$ as the eigenvectors of matrix $S$ corresponding to eigenvalues $\lambda$. Afterwards, we can extract factor vectors $V$ where the corresponding $V_i = X^T \frac{U_i}{\sqrt{\lambda_i}}$. Finally, the eigentriple of $\{\sqrt{\lambda_i}, U_i, V_i^T\}$ is formulated as a row-wise decomposition of matrix $X$ [@Golyandina2020]. The resulting eigentriple is then used to reconstruct the time-series by eigentriple grouping and diagonal averaging.
